---
title: "analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
library(tidyverse)

Clay had the longest long jump
```{r}
decathalon_data_long %>% 
  filter(sport_type == "Long.jump") %>% 
  arrange(desc(result)) %>% 
  head(1)

```
The mean Decastar result in the 100m was 11.17538
```{r}
decathalon_data_long %>% 
  filter(sport_type == "100m" & competition == "Decastar") %>% 
  summarise(decastar_result = mean(result))
  
```
The mean OlympicG result in the 100m was 10.91571
```{r}
decathalon_data_long %>% 
  filter(sport_type == "100m" & competition == "OlympicG") %>% 
  summarise(olympicg_result = mean(result))

```
Sebrle has the highest total points across both competitions
```{r}
decathalon_data_long %>% 
  filter(sport_type == "100m") %>% 
  group_by(name) %>% 
  summarise(total_points = sum(points)) %>% 
  arrange(desc(total_points)) %>% 
  head(1)
```
The top scorers for shot put in the Decastar competition were Yurkov, Sebrle and Karpov
```{r}
decathalon_data_long %>% 
  filter(sport_type == "Shot.put", competition == "Decastar") %>% 
  arrange(desc(result)) %>% 
  head(3)

  

```
The top scorers for shot put in the OlympicG competition were Macey, Sebrle and Karpov
```{r}
decathalon_data_long %>% 
  filter(sport_type == "Shot.put", competition == "OlympicG") %>% 
  arrange(desc(result)) %>% 
  head(3)
```
The average points for competitors who ran the 400m in more than 50 seconds is 7727.167
```{r}
decathalon_data_long %>% 
  filter(sport_type == "400m", result >= 50) %>% 
  summarise(mean_more_than_50 = mean(points))
```
The average points for competitors who ran the 400m in more than 50 seconds is 8120.483
```{r}
decathalon_data_long %>% 
  filter(sport_type == "400m", result < 50) %>% 
  summarise(mean_fewer_than_50 = mean(points))
```

