---
title: "Project Overview"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
ship_bird_data <- read_csv("../clean_data/ship_bird_data")
```



## Background Information

This dataset is about sightings of birds from ships between 1899 and 1990. It gives detailed information about the birds.

## Assumptions Made

* That some people might be interested in the sightings where there were no birds
* Kept the NA values as I decided there wasn't enough other information to impute 

## Challenges

* Confusion about loading here function
* Removing the extra information from the species columns
* Difficulty working out how to exclude birds that had been used more than once

## Data Cleaning Steps

* Removed columns without unique values
* Joined the pages together on the record_id column
* Removed extra information from the common names, scientific names and abbreviations 
* Reordered the columns

## Analysis

1. The wandering albatross sensu lato had the most sightings 
```{r}
ship_bird_data %>% 
  select(species_common_names, species_scientific_names, abbreviation) %>% 
  group_by(species_common_names) %>% 
  count(sightings = n()) %>% 
  arrange(desc(sightings)) %>% 
  head(1)
```

2. The short-tailed shearwater is the bird with the highest total count
```{r}
ship_bird_data %>%
  select(species_common_names, species_scientific_names, abbreviation, observations) %>%
  group_by(species_common_names) %>%
  tally(observations) %>% 
  arrange(desc(n)) %>% 
  head(1)
```

3. The wedge-tailed shearwater has the highest total above a latitude of -30
```{r}
ship_bird_data %>%
  select(species_common_names, species_scientific_names, abbreviation, observations, lat) %>%
  filter(lat > -30) %>% 
  group_by(species_common_names) %>%
  tally(observations) %>% 
  arrange(desc(n)) %>% 
  head(1)
```

4. 21 types of birds were only ever seen in groups of 1
```{r}
ship_bird_data %>% 
  select(species_common_names, species_scientific_names, abbreviation, observations)  %>% 
  group_by(species_common_names, species_scientific_names, abbreviation) %>% 
  summarise(maximum_observations = max(observations)) %>% 
  filter(maximum_observations == 1) %>% 
  nrow()  
```

5. 158 penguins were observed
```{r}
ship_bird_data %>% 
  select(species_common_names, species_scientific_names, abbreviation, observations) %>% 
  filter(str_detect(species_common_names, "penguin|Penguin")) %>% 
  tally(observations)
```
